---
title: "Comparing dendrograms from Datasets 1,2,3"
output: html_notebook
---

We want to compare the clusters identified from Datasets 1, 2, 3 to see how similar the results are.

To compare these dendrograms, we must first load required libraries and saved Rdata files.

```{r, message=F, error=F}
#library(renv) #remotes::install_github("rstudio/renv")
#renv::restore()
library(tidyverse)
library(here)
library(ggdendro)
library(cluster)
library(dendextend)
#renv::snapshot()
base::load(here::here("data", "Dataset1", "dataset1-ms1_dedro.RData"))          
dataset1_clust <- ms1_drugclust_svd
dataset1_clustboot <- ms1_drugclustboot_svd
dataset1_distance <- ms1_moduleTraitCor_svd

rm(ms1_drugclust_svd)
rm(ms1_drugclustboot_svd)
rm(ms1_moduleTraitCor_svd)

base::load(here::here("data", "Dataset2", "dataset2-msms_dedro.RData"))   

dataset2_clust <- msms_drugclust_svd
dataset2_clustboot <- msms_drugclustboot_svd
dataset2_distance <- msms_moduleTraitCor_svd

rm(msms_drugclust_svd)
rm(msms_drugclustboot_svd)
rm(msms_moduleTraitCor_svd)

base::load(here::here("data", "Dataset3", "dataset3-ms1_60dedro.RData"))   

dataset3_clust <- ms1_60drugclust_svd
dataset3_clustboot <- ms1_60drugclustboot_svd
dataset3_distance <- ms1_60moduleTraitCor_svd

rm(ms1_60drugclust_svd)
rm(ms1_60drugclustboot_svd)
rm(ms1_60moduleTraitCor_svd)
```

We have previously calculated the best `k` choice for each dendrogram, and so we will use this previous information where: 
- Dataset 1: k=5
- Dataset 2: k=3
- Dataset 3: k=4

# Dataset 1 vs Dataset 2

```{r}
pink <- "#DF737A"
blackcof <- "#2E1F27"
teal <- "#59C9A5"
yellow <- "#FFE66D"
blue <- "#3444DA"

clust_cols_D1 <- c("#DF737A", "#2E1F27", "#59C9A5","#FFE66D", "#3444DA")
clust_cols_D2 <-  c(pink, teal, blue)

dataset1_dend <- dataset1_clust %>% as.dendrogram() %>% color_branches(k=5, col=clust_cols_D1) %>%
    dendextend::set("branches_lwd", 3) 

dataset2_dend <- dataset2_clust %>% as.dendrogram() %>% color_branches(k=3, col=clust_cols_D2) %>%
    dendextend::set("branches_lwd", 3) 

compare_D1D2<- dendlist(dataset1_dend , dataset2_dend ) %>% untangle(method = "step2side")
compare_D1D2 %>% entanglement() #0.044


#pdf("../../4.mSystems/2.revisionv2/figs/tanglegramD1D2.pdf", width=6, height=5)
tanglegram(compare_D1D2, sort=F, common_subtrees_color_lines=F,highlight_branches_lwd = F, 
           highlight_distinct_edges=F, main="", main_left="Dataset 1", main_right="Dataset 2")
#dev.off()

```

```{r}
### cophenetic correlation value

## "The value can range between -1 to 1. With near 0 values meaning that the two trees are not statistically similar. For exact p-value one should result to a permutation test. One such option will be to permute over the labels of one tree many times, and calculating the distribution under the null hypothesis (keeping the trees topologies constant)." 
## Null hypothesis is correlation = 0 
## two tailed:  H1  cor != 0
## one tailed: H1 cor > 0 OR H1 cor < 0
factorial(15) # this is the top number of permutations...still quite a lot 

cor_coph <- cor_cophenetic(dataset1_dend, dataset2_dend)
set.seed(2248)
R <- 10000
cor_coph_results <- numeric(R)
dend_mixed <- dataset2_dend #shuffle dataset2_dend
for (i in 1:R){
  dend_mixed <- sample.dendrogram(dend_mixed, replace = F) #just swapping labels for a permutation p-value
  cor_coph_results[i] <- cor_cophenetic(dataset1_dend, dend_mixed) # compare to 
}

mean(abs(cor_coph_results)>=cor_coph)
sum(abs(cor_coph_results)>=cor_coph)
cor_coph_results <- cor_coph_results %>% as.data.frame()

cor_coph_round <- cor_coph %>% round(., digits=3)

(cor_coph_plot <-  ggplot(cor_coph_results, aes(x=.)) +
    geom_density(fill = "#59c9a5", colour = "#59c9a5", alpha = 0.8) +
    geom_vline(xintercept=1, size = 1, linetype = "dashed", colour="#2e1f27") +
    geom_vline(xintercept=0, size = 1, linetype = "dashed", colour="#2e1f27") +
    geom_vline(xintercept=cor_coph, size=1.2, colour="#df737a") +
    annotate("text", x=0.3, y=4, label= paste0("r = ",cor_coph_round) ) +
    ylab("Density") +
    xlab("Cophenetic correlation coefficent") + 
    theme_bw())
#ggsave(file="../../4.mSystems/2.revisionv2/figs/corD1D2.pdf", width=6, height=4, units="in")
```

# Dataset 2 vs 3

```{r}
#clust_cols_D3 <- c(pink, yellow, blue,teal)
clust_cols_D3 <- c(yellow, blue, pink,teal)
dataset3_dend <- dataset3_clust %>% as.dendrogram() %>% color_branches(k=4, col=clust_cols_D3) %>%
    dendextend::set("branches_lwd", 3) 
set.seed(123)
compare_D2D3<- dendlist(dataset2_dend , dataset3_dend ) %>% 
    untangle("random", R=5) %>%
    untangle(method = "step2side")
compare_D2D3 %>% entanglement() #0.347


#pdf("../../4.mSystems/2.revisionv2/figs/tanglegramD2D3.pdf", width=6, height=5)
tanglegram(compare_D2D3, sort=F, common_subtrees_color_lines=F,highlight_branches_lwd = F, 
           highlight_distinct_edges=F, main="", main_left="Dataset 2", main_right="Dataset 3")
#dev.off()
```

```{r}
### cophenetic correlation value

## "The value can range between -1 to 1. With near 0 values meaning that the two trees are not statistically similar. For exact p-value one should result to a permutation test. One such option will be to permute over the labels of one tree many times, and calculating the distribution under the null hypothesis (keeping the trees topologies constant)." 
## Null hypothesis is correlation = 0 
## two tailed:  H1  cor != 0
## one tailed: H1 cor > 0 OR H1 cor < 0
#factorial(15) # this is the top number of permutations...still quite a lot 

cor_coph <- cor_cophenetic(dataset2_dend, dataset3_dend)
set.seed(2248)
R <- 10000
cor_coph_results <- numeric(R)
dend_mixed <- dataset3_dend #shuffle dataset2_dend
for (i in 1:R){
  dend_mixed <- sample.dendrogram(dend_mixed, replace = F) #just swapping labels for a permutation p-value
  cor_coph_results[i] <- cor_cophenetic(dataset2_dend, dend_mixed) # compare to 
}

mean(abs(cor_coph_results)>=cor_coph)
sum(abs(cor_coph_results)>=cor_coph)
cor_coph_results <- cor_coph_results %>% as.data.frame()

cor_coph_round <- cor_coph %>% round(., digits=3)

(cor_coph_plot <-  ggplot(cor_coph_results, aes(x=.)) +
    geom_density(fill = "#59c9a5", colour = "#59c9a5", alpha = 0.8) +
    geom_vline(xintercept=1, size = 1, linetype = "dashed", colour="#2e1f27") +
    geom_vline(xintercept=0, size = 1, linetype = "dashed", colour="#2e1f27") +
    geom_vline(xintercept=cor_coph, size=1.2, colour="#df737a") +
    annotate("text", x=0.3, y=4, label= paste0("r = ",cor_coph_round) ) +
    ylab("Density") +
    xlab("Cophenetic correlation coefficent") + 
    theme_bw())
#ggsave(file="../../4.mSystems/2.revisionv2/figs/corD2D3.pdf", width=6, height=4, units="in")
```



# Dataset 1 vs Dataset 3

```{r}

compare_D1D3<- dendlist(dataset1_dend , dataset3_dend ) %>% untangle(method = "step2side")
compare_D1D3 %>% entanglement() #0.314


#("../../4.mSystems/2.revisionv2/figs/tanglegramD1D3.pdf", width=6, height=5)
tanglegram(compare_D1D3, sort=F, common_subtrees_color_lines=F,highlight_branches_lwd = F, 
           highlight_distinct_edges=F, main="", main_left="Dataset 1", main_right="Dataset 3")
#dev.off()

```

```{r}
### cophenetic correlation value

## "The value can range between -1 to 1. With near 0 values meaning that the two trees are not statistically similar. For exact p-value one should result to a permutation test. One such option will be to permute over the labels of one tree many times, and calculating the distribution under the null hypothesis (keeping the trees topologies constant)." 
## Null hypothesis is correlation = 0 
## two tailed:  H1  cor != 0
## one tailed: H1 cor > 0 OR H1 cor < 0
factorial(15) # this is the top number of permutations...still quite a lot 

cor_coph <- cor_cophenetic(dataset1_dend, dataset3_dend)
set.seed(2248)
R <- 10000
cor_coph_results <- numeric(R)
dend_mixed <- dataset3_dend #shuffle dataset2_dend
for (i in 1:R){
  dend_mixed <- sample.dendrogram(dend_mixed, replace = F) #just swapping labels for a permutation p-value
  cor_coph_results[i] <- cor_cophenetic(dataset1_dend, dend_mixed) # compare to 
}

mean(abs(cor_coph_results)>=cor_coph)
sum(abs(cor_coph_results)>=cor_coph)
cor_coph_results <- cor_coph_results %>% as.data.frame()

cor_coph_round <- cor_coph %>% round(., digits=3)

(cor_coph_plot <-  ggplot(cor_coph_results, aes(x=.)) +
    geom_density(fill = "#59c9a5", colour = "#59c9a5", alpha = 0.8) +
    geom_vline(xintercept=1, size = 1, linetype = "dashed", colour="#2e1f27") +
    geom_vline(xintercept=0, size = 1, linetype = "dashed", colour="#2e1f27") +
    geom_vline(xintercept=cor_coph, size=1.2, colour="#df737a") +
    annotate("text", x=0.3, y=4, label= paste0("r = ",cor_coph_round) ) +
    ylab("Density") +
    xlab("Cophenetic correlation coefficent") + 
    theme_bw())
#ggsave(file="../../4.mSystems/2.revisionv2/figs/corD2D3.pdf", width=6, height=4, units="in")
```